Towards Pandemic-Scale Ancestral Recombination Graphs of SARS-CoV-2

### 2023-06-02 paper.tex
Remove spaces

\section{Discussion}
% PARA 1. SARS-CoV-2 is still important, and the vast hordes of data provide an
% unprecedented opportunity to study the evolution of a virus. Even if the
% public health emergency is officially over, it doesn't mean it's not coming
% back and we should be prepared. Recombination is now a vital factor, and
% present a way in which it can be incorporated. Our results are accurate.
Although the COVID-19 pandemic is no longer considered a global emergency
by the WHO, the prevalence of SARS-CoV-2 continues to be high worldwide. This fuels
proliferation of many variants, with more than 600 Pango-designated lineages
circulating globally in the last three months (January to March, 2023;
https://gisaid.org/; access on March 27, 2023). High prevalence increases the risk of coinfection, providing opportunities for new phenotypically
distinct recombinants to emerge and spread. Phylogenetic approaches have
been central to pandemic responses thus far \citep{Attwood2022-ab}.
% 1. Form the basis of variant classifications (UShER paper),
% 2. Global context. In 2020, Nextstrain provided situation reports that were interactive, phylogenetically-driven visualizations of SARS-CoV-2 characteristics such as transmission https://nextstrain.org/ncov-sit-reps/
% 3. The WHO's recommendations on Omicron asked countries to "report initial cases/clusters associated with VOC infection to WHO". Phylogenetics being one way to identify those crucial early clusters. https://www.who.int/news/item/26-11-2021-classification-of-omicron-(b.1.1.529)-sars-cov-2-variant-of-concern
However, with the rise to high frequency of recombinant lineages \citep[e.g. XBB; ][]{Tamura2023-ab},
it is imperative that these methods are updated to incorporate the
effects of recombination, so that future public health interventions are
not based on incomplete and potentially biased evolutionary models.
Here we have introduced the first method to infer an evolutionary
history that jointly estimates genealogies with both mutation and
recombination at pandemic scale and illustrated how this single
structure accurately captures results derived by many different means.

>>>>

\section{Discussion}
% PARA 1. SARS-CoV-2 is still important, and the vast hordes of data provide an
% unprecedented opportunity to study the evolution of a virus. Even if the
% public health emergency is officially over, it doesn't mean it's not coming
% back and we should be prepared. Recombination is now a vital factor, and
% present a way in which it can be incorporated. Our results are accurate.
Although the COVID-19 pandemic is no longer considered a global emergency
by the WHO, the prevalence of SARS-CoV-2 continues to be high worldwide. This fuels
proliferation of many variants, with more than 600 Pango-designated lineages
circulating globally in the last three months (January to March, 2023;
\url{https://gisaid.org/}; accessed on 2023-03-27).
High prevalence increases the risk of coinfection,
providing opportunities for new phenotypically
distinct recombinants to emerge and spread. Phylogenetic approaches have
been central to pandemic responses thus
far~\citep{Attwood2022-ab,Bloom2023-fitness,Abbas2022-reconstruction,
Mclaughlin2022-genomic}.
However, with the rise to high frequency of recombinant
lineages \citep[e.g.\ XBB;][]{Tamura2023-ab},
it is imperative that these methods are updated to incorporate the
effects of recombination, so that future public health interventions are
not based on incomplete and potentially biased evolutionary models.
Here we have introduced the first method to infer an evolutionary
history that jointly estimates genealogies with both mutation and
recombination at pandemic scale and illustrated how this single
structure accurately captures results derived by many different means.
